`{peacesciencer}`:  A User’s  Guide  for  Quantitative  Peace Science in R

### 2021-06-10 user-guide/doc/ms.tex
meh

This data feature proceeds in the following fashion. The next section expands what need this package fills for peace scientists in the present. Afterward, it provides an overview of what is included in \texttt{\{peacesciencer\}} to help researchers more quickly conduct the kind of quantitative research they want. Thereafter, it provides a tutorial on how to install and best use \texttt{\{peacesciencer\}} in the R programming language. A more comprehensive tutorial follows, showing how \texttt{\{peacesciencer\}} already has a suite of data and functions that can allow for effective replications of a ``dangerous dyads'' type analysis (Bremer 1992) and standard state-year analyses of civil conflict onset (e.g. Fearon and Laitin 2003). This feature concludes with a discussion of what else \texttt{\{peacesciencer\}} is capable of doing in future updates and how this R package can inform more reasoned design decisions for researchers in peace science.

@@

\texttt{\{peacesciencer\}} is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch. NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars (Bennett, Poast, and Stam 2019). NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade (Bennett and Stam 2000). No matter, \texttt{\{peacesciencer\}} is motivated by the following ideals that led to its creation.

@@

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer (c.f. Bowers and Voors 2016). This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that \texttt{\{peacesciencer\}} uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet (Arel-Bundock 2021b), scrape \emph{any} type of data from the internet (Boehmke 2016), and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero (Arel-Bundock 2021a). \texttt{\{peacesciencer\}} embraces this. This R programming language package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

Fifth, statistical software should be kind to graduate students in peace science with the aforementioned point in mind. Graduate students just beginning their career in peace science scholarship are tasked to learn the substantive material, the characteristics of the underlying data in wide use in the community, and the quantitative methods and statistical software available to them, effectively all at the same time. A student interested in civil conflict scholarship will need to learn the UCDP-PRIO armed conflict data in some detail at the same time they learn about the theoretical arguments that informed the creation of the data, likely at the same time they take quantitative methods courses that show them how to load and explore data in a choseen statistical software. At the same time that creating data is increasingly blurred with coding and processing data, students are simultaneously learning to explore data for themselves as they read about the analyses of the data. \texttt{\{peacesciencer\}} aims to make this process easier for graduate students using the features of the R programming language. Graduate students can use \texttt{\{peacesciencer\}} to learn about peace science data and computer programming in R at the same time.

@@

\texttt{\{peacesciencer\}} comes with a battery of built-in functions for generating the most widespread forms of peace science data. Each function uses raw or pre-processed data included in \texttt{\{peacesciencer\}}. For example, \texttt{create\_dyadyears()} transforms the raw state system membership data for either the Correlates of War or Gleditsch-Ward system into the full universe of post-1816 dyad-years. \texttt{add\_gml\_mids()} uses a dyadic dispute-year version of the MID data offered by Gibler, Miller, and Little (2016), which, as the \texttt{data-raw} directory on the project's Github shows, is itself derived from the dispute-level and participant-level versions of the data. \texttt{\{peacesciencer\}} also has some data innovations. It has an original data set on capital-to-capital distances, calculating for distance in kilometers between state capitals using the ``Vincenty'' method (i.e.~``as the crow flies'') that also accounts for instances when capitals moved (e.g.~Brazil in 1960, Burundi in 2018). It also features an innovation of democracy data, providing reasonable estimates of democracy using the Marquez (2016) of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve, and Melton 2010).\footnote{A vignette on the \texttt{\{peacesciencer\}} website explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.} As of writing, \texttt{\{peacesciencer\}} comes with the following functionality.

@@

\texttt{\{peacesciencer\}}'s coverage focuses mostly on data that are released as standalone data sets for download, especially those in the Correlates of War or Gleditsch-Ward ecosystem of data. Data that can be obtained from a stable advanced programming interface---like the World Bank, for example---can be obtained through those other means (e.g. Arel-Bundock 2021b). Its coverage will assuredly expand with new additions of interest to the peace science community.

@@

\texttt{\{peacesciencer\}}'s functions work out of the box, though it works best with one additional package. This is \texttt{\{tidyverse\}}, itself a set of packages that share a common form and design (Wickham and Grolemund 2017). \texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}} but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality. A user can open an R session and install both packages as follows.

@@

The next step in the \texttt{\{peacesciencer\}} ``pipe''-based workflow will also be among the biggest time-savers for researchers using it. Researchers who want to understand, say, the correlates of inter-state conflict onset almost always do the following. They will add the dependent variables from an inter-state conflict data set, like the Gibler-Miller-Little (GML) dispute data set (Gibler, Miller, and Little 2016). They will add several indicators to the dyad-year data that are well-traveled in conflict scholarship, including indicators of things like joint alliance contracts and the contiguity relationship of both states in the dyad. \texttt{\{peacesciencer\}} has great coverage for a lot of these data and can even allow almost full-scale replications with the data it offers.

@@

After creating the base data they want, the user can add data to it with the ``pipe'' operator. The \texttt{\{tidyverse\}} represents this pipe operator as \texttt{\%\textgreater{}\%}. Ending each line with this pipe passes forward an object (here: data) into the next function on the next line. Put in other words, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)\ \%\textgreater{}\%\ filter\_prd()} is a command in \texttt{\{tidyverse\}} and \texttt{\{peacesciencer\}} to create all non-directed dyad-years from 1816 to 2016 and then subset those data to only the politically relevant dyad-years. Because politically relevant dyads depend on information about the major power status for both states as well as their contiguity relationship (Lemke and Reed 2001), the function adds in those data (i.e.~quietly executing \texttt{add\_contiguity()} and \texttt{add\_cow\_majors()}) before subsetting the data to just the dyads that have some kind of contiguity relationship and/or have a major power in it.

@@

Next, the pipe-based workflow continues with \texttt{add\_gml\_mids(keep\ =\ NULL)}, which will add information about whether there was an ongoing dispute in the dyad (and whether it was a dispute onset) from the GML dispute data.\footnote{\texttt{keep\ =\ NULL} is optional and removes supporting information about dyadic dispute-participants (e.g.~dispute-level hostility, dispute-level outcome) in order to reduce the size of the data set.} \texttt{add\_peace\_years()} will calculate peace years between ongoing disputes in these politically relevant non-directed dyads.\footnote{\texttt{add\_peace\_years()} will only calculate the peace years---titled, in this case, \texttt{gmlmidspell}---and will leave the temporal dependence adjustment to the taste of the researcher. Importantly, I do not recommend manually creating splines or square/cube terms because it creates more problems in adjusting for temporal dependence in model predictions. In a regression formula in R, you can specify the D. B. Carter and Signorino (2010) approach as \texttt{...\ +\ gmlmidspell\ +\ I(gmlmidspell\^{}2)\ +\ I(gmlmidspell\^{}3)}. The Beck, Katz, and Tucker (1998) cubic splines approach is \texttt{...\ +\ splines::bs(gmlmidspell,\ 4)}. This function includes the spell and three splines (hence the 4 in the command).} \texttt{add\_nmc()} will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity. \texttt{add\_democracy()} will add information about the level of democracy for both states in the dyad using three different indicators. Researchers can select which one they prefer from these indicators. \texttt{add\_cow\_alliance()} will add information about whether there was an alliance in the dyad-year by way of whether ther was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge. Finally, \texttt{add\_sdp\_gdp()} will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by Anders, Fariss, and Markowitz (2020).

@@

\texttt{\{peacesciencer\}} is also capable of creating data for replications of standard civil conflict analyses. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's (2003) well-cited analysis of civil conflict onset using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with \texttt{create\_stateyears(system\ =\ \textquotesingle{}gw\textquotesingle{})}, creating the full universe of Gleditsch-Ward state years from 1816 to 2020. Whereas the UCDP data included in \texttt{\{peacesciencer\}} are from 1946 to 2019, we can subset (``filter'') the data to just those years with \texttt{filter(year\ \%in\%\ c(1946,\ 2019))}. Next, we can use the \texttt{add\_ucdp\_acd()} function to return information about ongoing UCDP conflicts and onsets for these states. \texttt{add\_ucdp\_acd()} takes three arguments: \texttt{type}, \texttt{issue}, and \texttt{only\_wars}. \texttt{type} is an optional argument for the type of armed conflicts for which the researcher wants information. Options include ``extrasystemic,'' ``interstate,'' ``intrastate,'' and ``II'' (short for ``internationalized intrastate''). If no \texttt{type} is specified, the function returns information about ongoing disputes and onsets for all states for all tyeps of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g.~\texttt{type\ =\ c("intrastate",\ "II")}). \texttt{issue} is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict. Options include ``territory,'' ``government,'' and ``both.'' If no \texttt{issue} is specified, the function returns information for all conflicts irregarding the particular issue. \texttt{only\_wars} is an argument that subsets the data to just those with the intensity levels of ``war'' when \texttt{only\_wars\ =\ TRUE}. The argument defaults to \texttt{FALSE}, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, \texttt{add\_ucdp\_acd(type\ =\ "intrastate",\ only\_wars\ =\ FALSE))} returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.\footnote{The function also returns some background information of interest to the researcher, including the maximum intensity observed and the IDs associated with all ongoing conflicts in the state that year.}

@@

Finally, we can add some covariates of interest to these data. \texttt{add\_peace\_years()} calculates peace spells between ongoing conflicts in the data generated by \texttt{add\_ucdp\_acd()}. \texttt{add\_democracy()} adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' (2016) extension of Pemstein et al's (2010) Unified Democracy Scores). \texttt{add\_creg\_fractionalization()} adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois. \texttt{add\_sdp\_gdp()} will add information about a state's estimated GDP, population, and GDP per capita from the Anders, Fariss, and Markowitz (2020) simulations. Finally, \texttt{add\_rugged\_terrain()} provide two estimates of the ruggedness of a state's terrain. The first is the terrain ruggedness index calculated by Nunn and Puga (2012) and the second is the Gibler and Miller (2014) extension of the natural logged percentage of the state that is mountainous (originally calculated by Fearon and Laitin (2003)). At the end of the pipe, the data returned by \texttt{\{peacesciencer\}} is assigned to an object minimally called \texttt{Data}.

@@

This is not to say \texttt{\{peacesciencer\}} does everything, but \texttt{\{peacesciencer\}} can only evolve and expand on what it already does well. For example, \texttt{\{peacesciencer\}} does not right now handle \emph{k}-adic data (c.f. Poast 2010). However, the nature of \emph{k}-ads do not make for necessarily problematic data merges and should not be difficult to add for researchers who want it in future updates. Likewise, \texttt{\{peacesciencer\}} contains the Archigos leader data (Goemans, Gleditsch, and Chiozza 2009) and is able to transform that into interesting data at the dyad-level and state-level, but doing a leader-level analysis with information about leader willingness to use force (e.g. J. Carter and Smith 2020) is not right now available. However, these are possible expansions on what \texttt{\{peacesciencer\}} already does well. Thus, this package can only evolve to meet new analytical demands for the peace science community. Users are free to request new features as ``issues'' on the project's Github.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-andersetal2020bbgb}{}%
Anders, Therese, Christopher J. Fariss, and Jonathan N. Markowitz. 2020. {``Bread Before Guns or Butter: Introducing Surplus Domestic Product (SDP).''} \emph{International Studies Quarterly} 64 (2): 392--405.

\leavevmode\hypertarget{ref-arelbundock2021m}{}%
Arel-Bundock, Vincent. 2021a. \emph{Modelsummary: Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready}. \url{https://CRAN.R-project.org/package=modelsummary}.

\leavevmode\hypertarget{ref-arelbundock2021w}{}%
---------. 2021b. \emph{WDI: World Development Indicators and Other World Bank Data}. \url{https://CRAN.R-project.org/package=WDI}.

\leavevmode\hypertarget{ref-becketal1998tts}{}%
Beck, Nathaniel, Jonathan N. Katz, and Richard Tucker. 1998. {``Taking Time Seriously: Time-Series-Cross-Section Analysis with a Binary Dependent Variable.''} \emph{American Journal of Political Science} 42 (4): 1260--88.

\leavevmode\hypertarget{ref-bennettetal2019n}{}%
Bennett, D. Scott, Paul Poast, and Allan C. Stam. 2019. {``NewGene: An Introduction for Users.''} \emph{Journal of Conflict Resolution} 63 (6): 1579--92.

\leavevmode\hypertarget{ref-bennettstam2000e}{}%
Bennett, D. Scott, and Allan Stam. 2000. {``EUGene: A Conceptual Manual.''} \emph{International Interactions} 26 (2): 179--204.

\leavevmode\hypertarget{ref-boehmke2016dwr}{}%
Boehmke, Bradley C. 2016. \emph{Data Wrangling with {R}}. Springer.

\leavevmode\hypertarget{ref-bowersvoors2016hiyr}{}%
Bowers, Jake, and Maarten Voors. 2016. {``How to Improve Your Relationship with Your Future Self.''} \emph{Revista de Ciencia Politica} 36 (3): 829--48.

\leavevmode\hypertarget{ref-bremer1992ddc}{}%
Bremer, Stuart A. 1992. {``Dangerous Dyads: Conditions Affecting the Likelihood of Interstate War, 1816-1965.''} \emph{Journal of Conflict Resolution} 36 (2): 309--41.

\leavevmode\hypertarget{ref-cartersignorino2010bf}{}%
Carter, David B., and Curtis S. Signorino. 2010. {``Back to the Future: Modeling Time Dependence in Binary Data.''} \emph{Political Analysis} 18 (3): 271--92.

\leavevmode\hypertarget{ref-cartersmith2020fml}{}%
Carter, Jeff, and Charles E. Smith. 2020. {``A Framework for Measuring Leaders' Willingness to Use Force.''} \emph{American Political Science Review} 114 (4): 1352--58.

\leavevmode\hypertarget{ref-fearonlaitin2003eicw}{}%
Fearon, James D., and David D. Laitin. 2003. {``Ethnicity, Insurgency, and Civil War.''} \emph{American Political Science Review} 97 (1): 75--90.

\leavevmode\hypertarget{ref-giblermiller2014etts}{}%
Gibler, Douglas M., and Steven V. Miller. 2014. {``External Territorial Threat, State Capacity, and Civil War.''} \emph{Journal of Peace Research} 51 (5): 634--46.

\leavevmode\hypertarget{ref-gibleretal2016amid}{}%
Gibler, Douglas M., Steven V. Miller, and Erin K. Little. 2016. {``An Analysis of the {M}ilitarized {I}nterstate {D}ispute {(MID)} Dataset, 1816-2001.''} \emph{International Studies Quarterly} 60 (4): 719--30.

\leavevmode\hypertarget{ref-goemansetal2009ia}{}%
Goemans, Henk E., Kristian Skrede Gleditsch, and Giacomo Chiozza. 2009. {``Introducing {A}rchigos: A Dataset on Political Leaders.''} \emph{Journal of Peace Research} 46 (2): 269--83.

\leavevmode\hypertarget{ref-lemkereed2001rprd}{}%
Lemke, Douglas, and William Reed. 2001. {``The Relevance of Politically Relevant Dyads.''} \emph{Journal of Conflict Resolution} 45 (1): 126--44.

\leavevmode\hypertarget{ref-marquez2016qme}{}%
Marquez, Xavier. 2016. {``A Quick Method for Extending the {U}nified {D}emocracy {S}cores.''}

\leavevmode\hypertarget{ref-nunnpuga2012r}{}%
Nunn, Nathan, and Diego Puga. 2012. {``Ruggedness: The Blessing of Bad Geography in {A}frica.''} \emph{Review of Economics and Statistics} 94 (1): 20--36.

\leavevmode\hypertarget{ref-pemsteinetal2010dc}{}%
Pemstein, Daniel, Stephen A. Meserve, and James Melton. 2010. {``Democratic Compromise: A Latent Variable Analysis of Ten Measures of Regime Type.''} \emph{Political Analysis} 18 (4): 426--49.

>>>>

This data feature proceeds in the following fashion. The next section expands what need this package fills for peace scientists in the present. Afterward, it provides an overview of what is included in \texttt{\{peacesciencer\}} to help researchers more quickly conduct the kind of quantitative research they want. Thereafter, it provides a tutorial on how to install and best use \texttt{\{peacesciencer\}} in the R programming language. A more comprehensive tutorial follows, showing how \texttt{\{peacesciencer\}} already has a suite of data and functions that can allow for effective replications of a ``dangerous dyads'' type analysis (Bremer, 1992) and standard state-year analyses of civil conflict onset (e.g. Fearon \& Laitin, 2003). This feature concludes with a discussion of what else \texttt{\{peacesciencer\}} is capable of doing in future updates and how this R package can inform more reasoned design decisions for researchers in peace science.

@@

\texttt{\{peacesciencer\}} is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch. NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars (Bennett, Poast \& Stam, 2019). NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade (Bennett \& Stam, 2000). No matter, \texttt{\{peacesciencer\}} is motivated by the following ideals that led to its creation.

@@

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer (c.f. Bowers \& Voors, 2016). This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that \texttt{\{peacesciencer\}} uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet (Arel-Bundock, 2021a), scrape \emph{any} type of data from the internet (Boehmke, 2016), and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero (Arel-Bundock, 2021b). \texttt{\{peacesciencer\}} embraces this. This R programming language package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

Fifth, statistical software should be kind to graduate students in peace science with the aforementioned point in mind. Graduate students just beginning their career in peace science scholarship are tasked to learn the substantive material, the characteristics of the underlying data in wide use in the community, and the quantitative methods and statistical software available to them, effectively all at the same time. A student interested in civil conflict scholarship will need to learn the UCDP-PRIO armed conflict data in some detail at the same time they learn about the theoretical arguments that informed the creation of the data, likely at the same time they take quantitative methods courses that show them how to load and explore data in a chosen statistical software. At the same time that creating data is increasingly blurred with coding and processing data, students are simultaneously learning to explore data for themselves as they read about the analyses of the data. \texttt{\{peacesciencer\}} aims to make this process easier for graduate students using the features of the R programming language. Graduate students can use \texttt{\{peacesciencer\}} to learn about peace science data and computer programming in R at the same time.

@@

\texttt{\{peacesciencer\}} comes with a battery of built-in functions for generating the most widespread forms of peace science data. Each function uses raw or pre-processed data included in \texttt{\{peacesciencer\}}. For example, \texttt{create\_dyadyears()} transforms the raw state system membership data for either the Correlates of War or Gleditsch-Ward system into the full universe of post-1816 dyad-years. \texttt{add\_gml\_mids()} uses a dyadic dispute-year version of the MID data offered by Gibler, Miller \& Little (2016), which, as the \texttt{data-raw} directory on the project's Github shows, is itself derived from the dispute-level and participant-level versions of the data. \texttt{\{peacesciencer\}} also has some data innovations. It has an original data set on capital-to-capital distances, calculating for distance in kilometers between state capitals using the ``Vincenty'' method (i.e.~``as the crow flies'') that also accounts for instances when capitals moved (e.g.~Brazil in 1960, Burundi in 2018). It also features an innovation of democracy data, providing reasonable estimates of democracy using the Marquez (2016) of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve \& Melton, 2010).\footnote{A vignette on the \texttt{\{peacesciencer\}} website explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.} As of writing, \texttt{\{peacesciencer\}} comes with the following functionality.

@@

\texttt{\{peacesciencer\}}'s coverage focuses mostly on data that are released as standalone data sets for download, especially those in the Correlates of War or Gleditsch-Ward ecosystem of data. Data that can be obtained from a stable advanced programming interface---like the World Bank, for example---can be obtained through those other means (e.g. Arel-Bundock, 2021a). Its coverage will assuredly expand with new additions of interest to the peace science community.

@@

\texttt{\{peacesciencer\}}'s functions work out of the box, though it works best with one additional package. This is \texttt{\{tidyverse\}}, itself a set of packages that share a common form and design (Wickham \& Grolemund, 2017). \texttt{\{peacesciencer\}} functions make considerable use of the component packages of \texttt{\{tidyverse\}} but installing and loading \texttt{\{tidyverse\}} will allow the researcher to make quicker use of \texttt{\{peacesciencer\}}'s functionality. A user can open an R session and install both packages as follows.

@@

The next step in the \texttt{\{peacesciencer\}} ``pipe''-based workflow will also be among the biggest time-savers for researchers using it. Researchers who want to understand, say, the correlates of inter-state conflict onset almost always do the following. They will add the dependent variables from an inter-state conflict data set, like the Gibler-Miller-Little (GML) dispute data set (Gibler, Miller \& Little, 2016). They will add several indicators to the dyad-year data that are well-traveled in conflict scholarship, including indicators of things like joint alliance contracts and the contiguity relationship of both states in the dyad. \texttt{\{peacesciencer\}} has great coverage for a lot of these data and can even allow almost full-scale replications with the data it offers.

@@

After creating the base data they want, the user can add data to it with the ``pipe'' operator. The \texttt{\{tidyverse\}} represents this pipe operator as \texttt{\%\textgreater{}\%}. Ending each line with this pipe passes forward an object (here: data) into the next function on the next line. Put in other words, \texttt{create\_dyadyears(directed\ =\ FALSE,\ mry\ =\ FALSE)\ \%\textgreater{}\%\ filter\_prd()} is a command in \texttt{\{tidyverse\}} and \texttt{\{peacesciencer\}} to create all non-directed dyad-years from 1816 to 2016 and then subset those data to only the politically relevant dyad-years. Because politically relevant dyads depend on information about the major power status for both states as well as their contiguity relationship (Lemke \& Reed, 2001), the function adds in those data (i.e.~quietly executing \texttt{add\_contiguity()} and \texttt{add\_cow\_majors()}) before subsetting the data to just the dyads that have some kind of contiguity relationship and/or have a major power in it.

@@

Next, the pipe-based workflow continues with \texttt{add\_gml\_mids(keep\ =\ NULL)}, which will add information about whether there was an ongoing dispute in the dyad (and whether it was a dispute onset) from the GML dispute data.\footnote{\texttt{keep\ =\ NULL} is optional and removes supporting information about dyadic dispute-participants (e.g.~dispute-level hostility, dispute-level outcome) in order to reduce the size of the data set.} \texttt{add\_peace\_years()} will calculate peace years between ongoing disputes in these politically relevant non-directed dyads.\footnote{\texttt{add\_peace\_years()} will only calculate the peace years---titled, in this case, \texttt{gmlmidspell}---and will leave the temporal dependence adjustment to the taste of the researcher. Importantly, I do not recommend manually creating splines or square/cube terms because it creates more problems in adjusting for temporal dependence in model predictions. In a regression formula in R, you can specify the Carter \& Signorino (2010) approach as \texttt{...\ +\ gmlmidspell\ +\ I(gmlmidspell\^{}2)\ +\ I(gmlmidspell\^{}3)}. The Beck, Katz \& Tucker (1998) cubic splines approach is \texttt{...\ +\ splines::bs(gmlmidspell,\ 4)}. This function includes the spell and three splines (hence the 4 in the command).} \texttt{add\_nmc()} will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity. \texttt{add\_democracy()} will add information about the level of democracy for both states in the dyad using three different indicators. Researchers can select which one they prefer from these indicators. \texttt{add\_cow\_alliance()} will add information about whether there was an alliance in the dyad-year by way of whether ther was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge. Finally, \texttt{add\_sdp\_gdp()} will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by Anders, Fariss \& Markowitz (2020).

@@

\texttt{\{peacesciencer\}} is also capable of creating data for replications of standard civil conflict analyses. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's (2003) well-cited analysis of civil conflict onset using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with \texttt{create\_stateyears(system\ =\ \textquotesingle{}gw\textquotesingle{})}, creating the full universe of Gleditsch-Ward state years from 1816 to 2020. Whereas the UCDP data included in \texttt{\{peacesciencer\}} are from 1946 to 2019, we can subset (``filter'') the data to just those years with \texttt{filter(year\ \%in\%\ c(1946,\ 2019))}. Next, we can use the \texttt{add\_ucdp\_acd()} function to return information about ongoing UCDP conflicts and onsets for these states. \texttt{add\_ucdp\_acd()} takes three arguments: \texttt{type}, \texttt{issue}, and \texttt{only\_wars}. \texttt{type} is an optional argument for the type of armed conflicts for which the researcher wants information. Options include ``extrasystemic'', ``interstate'', ``intrastate'', and ``II'' (short for ``internationalized intrastate''). If no \texttt{type} is specified, the function returns information about ongoing disputes and onsets for all states for all tyeps of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g.~\texttt{type\ =\ c("intrastate",\ "II")}). \texttt{issue} is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict. Options include ``territory'', ``government'', and ``both''. If no \texttt{issue} is specified, the function returns information for all conflicts irregarding the particular issue. \texttt{only\_wars} is an argument that subsets the data to just those with the intensity levels of ``war'' when \texttt{only\_wars\ =\ TRUE}. The argument defaults to \texttt{FALSE}, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, \texttt{add\_ucdp\_acd(type\ =\ "intrastate",\ only\_wars\ =\ FALSE))} returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.\footnote{The function also returns some background information of interest to the researcher, including the maximum intensity observed and the IDs associated with all ongoing conflicts in the state that year.}

@@

Finally, we can add some covariates of interest to these data. \texttt{add\_peace\_years()} calculates peace spells between ongoing conflicts in the data generated by \texttt{add\_ucdp\_acd()}. \texttt{add\_democracy()} adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' (2016) extension of Pemstein et al's (2010) Unified Democracy Scores). \texttt{add\_creg\_fractionalization()} adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois. \texttt{add\_sdp\_gdp()} will add information about a state's estimated GDP, population, and GDP per capita from the Anders, Fariss \& Markowitz (2020) simulations. Finally, \texttt{add\_rugged\_terrain()} provide two estimates of the ruggedness of a state's terrain. The first is the terrain ruggedness index calculated by Nunn \& Puga (2012) and the second is the Gibler \& Miller (2014) extension of the natural logged percentage of the state that is mountainous (originally calculated by Fearon \& Laitin (2003)). At the end of the pipe, the data returned by \texttt{\{peacesciencer\}} is assigned to an object minimally called \texttt{Data}.

@@

This is not to say \texttt{\{peacesciencer\}} does everything, but \texttt{\{peacesciencer\}} can only evolve and expand on what it already does well. For example, \texttt{\{peacesciencer\}} does not right now handle \emph{k}-adic data (c.f. Poast, 2010). However, the nature of \emph{k}-ads do not make for necessarily problematic data merges and should not be difficult to add for researchers who want it in future updates. Likewise, \texttt{\{peacesciencer\}} contains the Archigos leader data (Goemans, Gleditsch \& Chiozza, 2009) and is able to transform that into interesting data at the dyad-level and state-level, but doing a leader-level analysis with information about leader willingness to use force (e.g. Carter \& Smith, 2020) is not right now available. However, these are possible expansions on what \texttt{\{peacesciencer\}} already does well. Thus, this package can only evolve to meet new analytical demands for the peace science community. Users are free to request new features as ``issues'' on the project's Github.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-andersetal2020bbgb}{}%
Anders, Therese, Christopher J Fariss \& Jonathan N Markowitz (2020) Bread before guns or butter: Introducing surplus domestic product (SDP). \emph{International Studies Quarterly} 64(2): 392--405.

\leavevmode\hypertarget{ref-arelbundock2021m}{}%
Arel-Bundock, Vincent (2021b) \emph{Modelsummary: Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready} (\url{https://CRAN.R-project.org/package=modelsummary}).

\leavevmode\hypertarget{ref-arelbundock2021w}{}%
Arel-Bundock, Vincent (2021a) \emph{WDI: World Development Indicators and Other World Bank Data} (\url{https://CRAN.R-project.org/package=WDI}).

\leavevmode\hypertarget{ref-becketal1998tts}{}%
Beck, Nathaniel, Jonathan N Katz \& Richard Tucker (1998) Taking time seriously: Time-series-cross-section analysis with a binary dependent variable. \emph{American Journal of Political Science} 42(4): 1260--1288.

\leavevmode\hypertarget{ref-bennettetal2019n}{}%
Bennett, DScott, Paul Poast \& Allan C Stam (2019) NewGene: An introduction for users. \emph{Journal of Conflict Resolution} 63(6): 1579--1592.

\leavevmode\hypertarget{ref-bennettstam2000e}{}%
Bennett, DScott \& Allan Stam (2000) EUGene: A conceptual manual. \emph{International Interactions} 26(2): 179--204.

\leavevmode\hypertarget{ref-boehmke2016dwr}{}%
Boehmke, Bradley C (2016) \emph{Data Wrangling with {R}}. Springer.

\leavevmode\hypertarget{ref-bowersvoors2016hiyr}{}%
Bowers, Jake \& Maarten Voors (2016) How to improve your relationship with your future self. \emph{Revista de Ciencia Politica} 36(3): 829--848.

\leavevmode\hypertarget{ref-bremer1992ddc}{}%
Bremer, Stuart A (1992) Dangerous dyads: Conditions affecting the likelihood of interstate war, 1816-1965. \emph{Journal of Conflict Resolution} 36(2): 309--341.

\leavevmode\hypertarget{ref-cartersignorino2010bf}{}%
Carter, David B \& Curtis S Signorino (2010) Back to the future: Modeling time dependence in binary data. \emph{Political Analysis} 18(3): 271--292.

\leavevmode\hypertarget{ref-cartersmith2020fml}{}%
Carter, Jeff \& Charles E Smith (2020) A framework for measuring leaders' willingness to use force. \emph{American Political Science Review} 114(4): 1352--1358.

\leavevmode\hypertarget{ref-fearonlaitin2003eicw}{}%
Fearon, James D \& David D Laitin (2003) Ethnicity, insurgency, and civil war. \emph{American Political Science Review} 97(1): 75--90.

\leavevmode\hypertarget{ref-giblermiller2014etts}{}%
Gibler, Douglas M \& Steven V Miller (2014) External territorial threat, state capacity, and civil war. \emph{Journal of Peace Research} 51(5): 634--646.

\leavevmode\hypertarget{ref-gibleretal2016amid}{}%
Gibler, Douglas M, Steven V Miller \& Erin K Little (2016) An analysis of the {M}ilitarized {I}nterstate {D}ispute {(MID)} dataset, 1816-2001. \emph{International Studies Quarterly} 60(4): 719--730.

\leavevmode\hypertarget{ref-goemansetal2009ia}{}%
Goemans, Henk E, Kristian Skrede Gleditsch \& Giacomo Chiozza (2009) Introducing {A}rchigos: A dataset on political leaders. \emph{Journal of Peace Research} 46(2): 269--83.

\leavevmode\hypertarget{ref-lemkereed2001rprd}{}%
Lemke, Douglas \& William Reed (2001) The relevance of politically relevant dyads. \emph{Journal of Conflict Resolution} 45(1): 126--144.

\leavevmode\hypertarget{ref-marquez2016qme}{}%
Marquez, Xavier (2016) A quick method for extending the {U}nified {D}emocracy {S}cores.

\leavevmode\hypertarget{ref-nunnpuga2012r}{}%
Nunn, Nathan \& Diego Puga (2012) Ruggedness: The blessing of bad geography in {A}frica. \emph{Review of Economics and Statistics} 94(1): 20--36.

\leavevmode\hypertarget{ref-pemsteinetal2010dc}{}%
Pemstein, Daniel, Stephen A Meserve \& James Melton (2010) Democratic compromise: A latent variable analysis of ten measures of regime type. \emph{Political Analysis} 18(4): 426--449.

\leavevmode\hypertarget{ref-poast2010mdd}{}%
Poast, Paul (2010) (Mis)using dyadic data to analyze multilateral events. \emph{Political Analysis} 18(4): 403--425.

\leavevmode\hypertarget{ref-wickhamgrolemund2017rds}{}%
Wickham, Hadley \& Garrett Grolemund (2017) \emph{{R} for Data Science}. Sebastopol, CA: O'Reilly Media, Inc.
